The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-02-12 paper/paper.tex
more in data section

\begin{itemize}
    \item CosmoDC2 - description
    \item CosmoDC2 - discussion
    \item Noise properties
    \item Pre-selection 
    \item Metacal as a limitation
\end{itemize}

@@

\subsection{ {\sc ENSEMBLE} }

@@

\section{Overall results}
The methods described above were run on a data set, previously unseen by entrants but drawn
from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
for each method.  In total, each method was run on 16 scenarios, each combination of: griz vs riz,
Buzzard vs CosmoDC2, and 3,5,7, and 9 bins requested.  Nine metrics were calculated for each
scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}, the $w_0-w_a$ 
DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).

>>>>

We run challenge entries on two data sets, CosmoDC2 and Buzzard, each of which is split into training, 
validation, and testing components.  In each case participants were supplied with the training and 
validation data, and the testing data was kept secret for final evaluation. In each case the three
data sets were random sub-selections of the full data sets, with the training and testing sets 25\% of 
the full data each and the validation 50\%. In practice many entrants trained on a subset of
this, so to make the final evaluation phase tractable we trained on 1M objects for both data sets.

Each of these data set come from cosmological simulations which provide truth values of galaxy magnitudes
and sizes.  We simulate and add noise to the objects as described below.

\subsubsection{CosmoDC2}

The CosmoDC2 simulation was designed to support the Dark Energy Science
Collaboration and is described in detail in \citet{cosmodc2}.  It covered 440
deg${}^2$ of sky, and used the dark matter particles from the Outer Rim
simulations \citep{outer_rim}.  The UniverseMachine \citep{universe_machine}
simulations were then used, in a combination of empirical and semi-analytic
methods, to assign galaxies with a limited set of specified properties to halos.

These objects were then matched to outputs of the Galacticus model
\citep{galacticus}, which generated complete galaxy properties and which was run
on an additional DESC simulation, AlphaQ, and included all the LSST bands.  The
simulation is complete to $r=28$, and contains objects up to redshift 3.  We
include the ultra-faint galaxies in the simulation, not assigned to a specific
halo.

After the post-processing and selections described below, and a cut to a specific geographic region, 
we used around 36M objects from the CosmoDC2 simulations.

One limitation of the CosmoDC2 catalogs we used, found during the challenge,
was that the matching to Galacticus led to only a limited number of SEDs being
used at high redshifts, and thus to many galaxies in the simulations sharing
similar characteristics at these redshifts, such as colour-colour relations.
It was unknown whether this limitation would have any practical impact on the challenge,
and was one of the reasons for adpoting the additional Buzzard catalog.

\subsubsection{Buzzard}

The Buzzard catalog was developed to support Dark Energy Survey Year 1 (DES-Y1)
data analysis, and is described in detail in \citet{buzzard}.  It used dark
matter simulations from L-GADGET2 \cite{lgadget} and then added galaxies using
the {\sc AddGals} algorithm \citet{addgals}, which matches a set of galaxy
property probability densities to data using sub-halo abundance matching. 
Buzzard galaxies extend to around redshift 2.3, and were shown to be a good
match (after appropriate selections) to DES-Y1 observable catalogs.  Magnitudes
using the LSST band passes were provided in the catalog.

After the post-processing and selections described below, and a cut to a
specific geographic region,  we used around 20M objects from the Buzzard
simulations.

\subsubsection{Post-processing}

We add noise to the truth values in the extra-galactic catalogs to simulate real observations.
In each case we simulate observations corresponding to the first year Rubin observations
using the DESC {\sc TXPipe} framework, following the methodology of \citet{ivezic_jones_lupton}
and assuming the numerical values for telescope and sky characteristics therein.

In both simulations we add two cuts to approximately simulate the selection used in a real
lensing catalog.  In both cases we apply a cut on the combined $riz$ signal to noise of
$S/N > 10$, and a size cut $T / T_\mathrm{psf} > 0.5$, where
$T$ is the trace $I_{xx} + I_{yy}$ of the moments matrix and measures the squared radius of the
galaxy compared to that of the PSF.

@@

@@

\subsection{Overall results}
The methods described above were run on a data set, previously unseen by entrants but drawn
from the same underlying population, of size 8.6M objects for CosmoDC2 and 5.4M objects for Buzzard.
A complete suite of the metrics described in Section \ref{sec:metrics} was run on each of the two
for each method.  In total, each method was run on 16 scenarios, each combination of: griz vs riz,
Buzzard vs CosmoDC2, and 3,5,7, and 9 bins requested.  Nine metrics were calculated for each
scenario: clustering, lensing, and 3x2pt for each of the SNR metric (Equation \ref{eq:snr}, the $w_0-w_a$ 
DETF FOM, and the $\Omega_c - \sigma_8$ FOM (Equation \ref{eq:fom}).
